The Birth of a Galaxy: Primordial Metal Enrichment and Stellar Populations

### 2011-11-18 ApJ/ms.tex
Adding OCI fellowship.
Addressing the final point in the referee report by rewriting a couple of paragraphs in the discussion and summary.  Updating preprint references.

Recall that the most massive halo ($M_{\rm tot} = 10^9 \Ms$) is
undergoing a starburst at $z=7$, translating into a relatively high
$M/L = 6$.  The lower mass galaxies show a spread in M/L between 30
and 3000, similar to the Sloan Digital Sky Survey (SDSS) observations
of local ultra faint dwarf (uFd) galaxies \citep[e.g.][]{Strigari08}.
This is suggestive that these low-mass ($M_{\rm tot} \lsim 10^8 \Ms$)
galaxies that cannot cool through hydrogen line cooling are the
progenitors of uFd galaxies \citep{Bovill11a, Bovill11b}.  In this
scenario, all of the star formation occurs before reionization and
was subsequently terminated by UV heating from external sources.

@@

We have focused on the birth of two galaxies prior to reionization
with a cosmological AMR radiation hydrodynamics simulation with
unprecedented spatial (1 pc comoving) and mass resolution (1800 \Ms).
With this resolution, we are able to follow the Pop III star formation
and radiative and chemical feedback effects in hundreds of minihalos
that contribute to the formation of a protogalaxy.  Supernovae from
Pop III stars provide the necessary heavy elements for the transition
to a Population II stellar population, which we have directly
simulated.  These two galaxies have a 10--15\% probability in
surviving as present-day ``fossil'' galaxies \citep{Gnedin06},
otherwise they will be incorporated into galactic stellar halos.  A
$z=8.55$ galaxy was recently spectroscopically confirmed that is
contained in an ionized bubble with radius 0.1--0.5 Mpc with an
uncharacterized population of galaxies contributing to the local
ionizing radiation field \citep{Lehnert10_z8.6}.  The galaxies
simulated here might represent this population and should be detected
with the upcoming \textit{James Webb Space Telescope}.

@@

In the case where this metallicity floor is less than the critical
metallicity, then the next SN will further enrich the ISM, solidifying
the transition to Pop II SF.  We conclude that it only takes one, at
most two, SNe from Pop III stars in the halo progenitors to complete
the transition to Population II \citep{Frebel10}.  However, this does
not prohibit more than two instances of Pop III star formation in halo
from happening because the stars can end in a direct collapse to a BH
with little or no metal ejecta, keeping the surrounding environment
pristine.  The question of whether the critical metallicity is
\tento{-6} or \tento{-3.5} $Z_\odot$ is most applicable to nearby
halos where the heavy elements mix slowly into dense cores as the
blastwave overtakes it.  Less than two percent of stars have [Z/H] $<
-3$ in both systems, consistent with observations of metal-poor stars
in the halo and dSphs \citep{Beers05, Battaglia11}.

@@

JHW thanks Renyue Cen, Amina Helmi, Marco Spaans, and Eline Tolstoy
for enlightening discussions.  JHW is partially supported by NASA
through Hubble Fellowship grant \#120-6370 awarded by the Space
Telescope Science Institute, which is operated by the Association of
Universities for Research in Astronomy, Inc., for NASA, under contract
NAS 5-26555.  MJT and MLN acknowledge partial support from NASA ATFP
grant NNX08AH26G.  Computational resources were provided by NASA/NCCS
award SMD-09-1439.  TA acknowledges financial support from the {\em
  Baden-W\"{u}rttemberg-Stiftung} under grant P-LS-SPII/18, the
Heidelberg Institut f\"ur Theoretische Studien. This work was
partially supported by NASA ATFP grant NNX08AH26G, NSF AST-0807312,
and NSF AST-1109243. This research has made use of NASA’s Astrophysics
Data System Bibliographic Services.  The majority of the analysis and
plots were done with \texttt{yt} \citep{yt_full_paper}.

>>>>

Recall that the most massive halo ($M_{\rm tot} = 10^9 \Ms$) is
undergoing a starburst at $z=7$, translating into a relatively high
$M/L = 6$.  The lower mass galaxies show a spread in M/L between 30
and 3000, similar to the Sloan Digital Sky Survey (SDSS) observations
of local ultra faint dwarf (uFd) galaxies \citep[e.g.][]{Strigari08}.
It has been suggested that these low-mass ($M_{\rm tot} \lsim 10^8
\Ms$) galaxies that cannot cool through hydrogen line cooling are the
progenitors of uFd galaxies \citep{Bovill11a, Bovill11b}, where all of
the star formation occurs before reionization and was subsequently
terminated by UV heating from external sources.  These galaxies have
stellar masses ranging from 210 \Ms~(SDSS J1058+2843) to $3.2 \times
10^5$ \Ms~(Dra) with most between $10^3$ and $10^4$
\Ms. \citep{Martin08_uFd}.  Because the total stellar population would
be sampled, on average, by several star particles, we cannot make
robust conclusions about the nature of uFd galaxies.

@@

We have focused on the birth of two galaxies prior to reionization
with a cosmological AMR radiation hydrodynamics simulation with very
high spatial (1 pc comoving) and mass resolution (1800 \Ms).  With
this resolution, we are able to follow the Pop III star formation and
radiative and chemical feedback effects in hundreds of minihalos that
contribute to the formation of a protogalaxy.  Supernovae from Pop III
stars provide the necessary heavy elements for the transition to a
Population II stellar population, which we have directly simulated.
The galaxies simulated here represent a population of uncharacterized
high-redshift dwarf galaxies, which should be detected with the
upcoming \textit{James Webb Space Telescope}, that should contribute a
significant fraction of the ionizing photons during reionization.
Recently, a $z=8.6$ spectroscopically confirmed galaxy was found to
reside in an \ion{H}{2} region that extends beyond 1 Mpc.  However,
using the SFRs and \lya~continuum as estimators, this galaxy produces
enough ionizing radiation to create an \ion{H}{2} region between $\sim
0.1(f_{\rm esc}/0.1)^{1/3}$ Mpc and $0.5(f_{\rm esc}/0.1)^{1/3}$ Mpc,
where $f_{\rm esc}$ is the fraction of ionizing radiation escaping
into the IGM \citep{Lehnert10_z8.6}.  This discrepancy could be caused
by undetected satellite dwarf galaxies, such ones presented here, that
contribute to local radiation field, or an ionizing radiation escape
fraction near unity \citep{Wise09, Razoumov10, Paardekooper11,
  Yajima11}.

@@

Building upon this idea of a lower metallicity floor, consider a
pre-galactic halo that has hosted a Pop III star, producing a SN.  The
majority of the gas is ejected into the IGM and re-accretes after
$\sim 10-50$ Myr \citep{Johnson07, Wise08_Gal}.  What happens if the
metallicity is not greater than the critical metallicity?  Massive Pop
III stars would form again in this halo; however, enriching the
surrounding IGM with Pop III SNe and ejecting material becomes more
difficult in halos with $M \gsim 10^7 \Ms$ \citep{Whalen08_SN}.  Thus
it is more likely for these subsequent Pop III stars to retain more of
its gas during star formation episodes and self-enrich its host halo
and less of the IGM.  This would accelerate the transition from Pop
III to II in more massive halos.

Using our result of the metallicity floor, we conclude that it only
takes one, at most two, SNe from Pop III stars in the halo progenitors
to complete the transition to Population II \citep{Frebel10}.
However, this does not prohibit more than two instances of Pop III
star formation in halo from happening because the stars can produce
BHs with little or no metal ejecta, keeping the surrounding
environment pristine.  Although we find that Pop III SNe create a
floor of $10^{-4} - 10^{-3} \zsun$, gas with $Z/\zsun < 10^{-4}$ can
still exists in nearby halos, where the metal-rich SN ejecta mix
slowly into dense cores as the blastwave overtakes it \citep{Cen08}.
These environments may be the host of true extremely metal-poor star
formation, such as the recently discovered halo star with all detected
metals below $10^{-4.2} \zsun$ \citep{Caffau11}.

@@

JHW thanks Renyue Cen, Amina Helmi, Marco Spaans, and Eline Tolstoy
for enlightening discussions.  We thank an anonymous referee for
helpful suggestions.  JHW is partially supported by NASA through
Hubble Fellowship grant \#120-6370 awarded by the Space Telescope
Science Institute, which is operated by the Association of
Universities for Research in Astronomy, Inc., for NASA, under contract
NAS 5-26555.  MJT and MLN acknowledge partial support from NASA ATFP
grant NNX08AH26G.  MJT~acknowledges support by the NSF CI TraCS
fellowship award OCI-1048505.  Computational resources were provided
by NASA/NCCS award SMD-09-1439.  TA acknowledges financial support
from the {\em Baden-W\"{u}rttemberg-Stiftung} under grant
P-LS-SPII/18, the Heidelberg Institut f\"ur Theoretische Studien. This
work was partially supported by NASA ATFP grant NNX08AH26G, NSF
AST-0807312, and NSF AST-1109243. This research has made use of NASA’s
Astrophysics Data System Bibliographic Services.  The majority of the
analysis and plots were done with \texttt{yt} \citep{yt_full_paper}.
